### 2024-06-14 manuscript.Rmd
make changes from ES comments

---
title: "Dissociation between vertebrae number and body size in Australian blindsnakes"
date: "Version `r Sys.Date()`"
author: ""
output: 
  # bookdown::html_document2:
  #   toc: true
  #   number_sections: FALSE
  bookdown::word_document2:
    reference_docx: [./custom-reference-doc.docx]
    number_sections: false
    fig_caption: yes
bibliography: [C:/Users/ST/Documents/repo/anu_thesis_texts/zotero_library.bib]
csl: evolution.csl
link-citations: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE,
                      dpi=300)
# knitr::opts_knit$set(root.dir=normalizePath("..")) #use grandparent folder working directory
# for including graphics we need full path with knitr
WD <- getwd()
```

```{r packages}
library(dplyr); library(readr); library(janitor); 
library(ape); library(phytools);
library(kableExtra)
```

<!-- **Title** TBD -->

**Authors**: Sarin Tiatragul^1^; Andrew Amey^2^; Scott Hocknull^2^; Scott Keogh^1^; Emma Sherratt^3^

^1^Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra 2601, ACT, Australia

^2^Queensland Museum, Hendra, QLD, 4011, Australia

^3^School of Biological Sciences, The University of Adelaide, Adelaide, SA 5005, Australia

**Corresponding author**: Sarin Tiatragul (sarin.tiatragul@anu.edu.au)

<!-- **Author contributions**:  -->

<!-- ST and ES conceptualised the study; ST collected morphology data;  -->
<!-- ST obtained X-rays from SAM, WAM, and MAGNT. SH and AM provided X-rays Queensland Museum specimens. -->
<!-- ST and ES analysed the data; ST prepared the figures; ST, SK and ES wrote the manuscript. -->

<!-- **Acknowledgements (including funding information)**: -->

<!-- We to thank museum curators and collection managers from the  -->
<!-- South Australian Museum  (R Foster, S Donnellan), Western Australian Museum (P Doughty), Queensland Museum (A Amey and P Couper), Australian National Wildlife Collection (T Haff, C Wilson, and L Joseph), Museum and Art Gallery of the Northern Territory (G Dally), Museums Victoria (J Sumner, J Melville, and R Aguilar), Florida Museum of Natural History (J Gray, CM Sheehy, and ZS Randall), and the Natural History Museum Berlin (K Mahlow and F Tillack).  -->
<!-- SK thanks the Australian Research Council for ongoing support. ST PhD scholarship is supported by the Australian Government Research Training Program Scholarship at the Australian National University.  -->

**Conflict of Interest statement**
The authors declare no Conflict of Interest.

**Data Availability Statement**
Upon acceptance, associated data will be archived on Zenodo/Dryad.

\newpage

# Abstract

Keywords: Scolecophidia, vertebrae, pleomerism, snakes,

\newpage

# Introduction

* Axial evolution an important aspect of the vertebrate body plan evolution – facilitates diversity and influences locomotion 
  <!-- Note the papers  that are about axial number relating to locomotion -->
@galisFast2014 - mammals
@vandammeSpeed2002 - lacertids
@yamadaAdaptive2009;@forkerBendy2021 - gobies 
@molnarMorphological2023 - chameleons

* Need a reference for mammals? Limited variation of a common theme that is constrained [@asherVariability2011;@naritaEvolution2005]
@liDivergent2023 found vertebral counts to differ significantly between major groups of mammals and within groups. 
* Reptiles, on the other hand, are more varied [@headEvolution2015]

* Elongate animals have special properties of vertebral number known as pleomerism, where body size and number of segments in vertebrae are correlated [@lindseyPleomerism1975]. 
Pleomerism has been observed in eels and fish [@wardEvolution2007;@wardDifferential2014;@reeceEvolutionary2013;], in squamates [@pollyTesting2001;@headDissociation2007;@headEvolution2015;@bergmannVertebral2012;@lindellEvolution1994;@cernanskyVertebral2019], in salamanders and caecilians [@bonettEvidence2017;@renousBody1989;@arntzenVertebral2015]

* Snakes are a good model to study axial evolution because they have a high number of vertebrae and a wide range of body size and shape.
* In elapid snakes [@sherrattVertebral2019;@sherrattMacroevolution2022;@hamptonInterspecific2019;@hamptonRegionalization2023]; colubrid snakes [@nash-hahnModification2024].

***[Specific background]***

Scolecophidians, often referred to in English as "blindsnakes", are characterised by their fossorial lifestyle, poor visual system, and highly specialised phenotypes adapted for a strictly fossorial existence [@mirallesMolecular2018; @strongDeconstructing2021; @gowerEyetranscriptome2021]. Scolecophidians comprise five paraphyletic families, including the "core scolecophidians" — a monophyletic clade consisting of Typhlopidae (425 spp), Leptotyphlopidae (143 spp), Gerrhopilidae (29 spp), and Xenotyphlopidae (1 spp) — and the paraphyletic Anomalepididae (20 spp), which is sister to all other "typical snakes" (Alethinophidia) [@mirallesMolecular2018; @burbrinkInterrogating2020; @uetzQuarter2021].

Although all blindsnakes are fossorial, body size and shape vary significantly among species (Tiatragul et al. unpublished; @shineNatural1990;@webbPreysize1993;@webbDietary2001;@ehmannFamily1993). Leptophlopids are generally thought to be more narrow-and-long compared to typhlopids, which are generally wide-and-robust for a given body length [@herrelBurrowing2021]. These differences in body size and proportions determine how much push force the snakes can generate when burrowing. Their morphological differences led to the idea that leptyphlopids generally use existing burrows while typhlopids create their own burrow.

***[Knowledge gap]***

The Australian blindsnake radiation comprise 47 described species [@tiatragulAustralia2023] with considerably diverse morphology from narrow-and-long (aspect ratio >= 100) to robust-and-wide (aspect ratio < 60). However, information about the relationship between body size and vertebrae number is unknown.

Among typhlopidae, gigantism have evolved multiple times but with different mechanisms [@headDissociation2007]. For instance, @headDissociation2007 found that some of the largest species of blindsnakes in the African continent (e.g. *Afrotyphlops schlegelii*) have fewer than expected precloacal vertebrae than expected for their size. This suggests that elongation in these species is achieved by elongating each vertebra rather than adding additional vertebrae.

In another predominantly burrowing clade, caecilians, the number of vertebrae is positively correlated with body size [@renousBody1989], suggesting pleomerism is a mechanism for burrowing taxa to diversity in the underground environment.

<!-- A big review included number of vertebrae [@szyndlarIllustrated2023] for many non-alethinophidian snakes, but  -->

Compared to other snakes, the vertebral column of fossorial snakes tend to feature a much higher proportion of pre- than post-cloacal segments [@pollyTesting2001]. This is reflected in the relatively short tails of blindsnakes (CITE).

***[What we do in this paper]***

In this study, we investigate the prevalence of pleomerism among Australian blindsnakes and test hypotheses that deviations from pleomerism may be explained by different body shapes or other ecological factors among Australian typhlopid snakes. The specific questions we addressed are: (1) How prevalent is pleomerism among Australian blindsnakes? (2) Can deviations from pleomerism be explained by body shape in a genus of Australian blindsnakes? (3) Does vertebral ratio vary with soil compactness or mean annual temperature?

<!-- Specifically, we test two predictions: -->
<!-- (1) narrow-and-long species develop from postnatal ontogenetic changes where vertebral numbers are not significantly different between narrow-and-long species compared to wide-and-robust species.  -->
<!-- (2) Narrow-and-long blindsnakes develop via changes in the embryonic and postnatal stages where narrow-and-long species have greater number of vertebrae than wide-and-robust species.  -->

<!-- To answer these questions we gathered data on the number of vertebrae across XX species of Australian blindsnakes.  -->

# Material and Methods

## Sampling

We examined 341 specimens representing 44 lineages *Anilios* blindsnakes including 42/47 formally described species [@uetzQuarter2021]. We additionally included two close relatives, *Ramphotyphlops multilineatus* and *Acutotyphlops subocularis* [@tiatragulAustralia2023]. Our sample size per species ranged from 1 to 25 specimens (median=7 specimens). We sourced specimens from South Australian Museum (SAMA), Western Australian Museum (WAM), Queensland Museum (QM), Museum and Art Gallery of the Northern Territory (MAGNT), and a private collection. Where possible, we chose relatively large non-fragile specimens from both males and females with well-preserved heads for X-ray. Sex for each specimen were determined by inspecting the gonads or assigned based on relative tail lengths following Tiatragul et al. (unpublished).

## Phylogeny

We extracted 37 lineages of Australian blindsnakes from the time-calibrated phylogeny of Australian blindsnakes and outgroups presented in [@tiatragulAustralia2023]. The subset phylogeny represents 35/47 formally described species of _Anilios_ along with two additional genetically distinct lineages: one eastern population of "_A. grypus_" and one eastern population of "_A. ligatus_" [@tiatragulAustralia2023]. This phylogeny was inferred using the shortcut coalescent method in ASTRAL III [@zhangASTRALIIIPolynomialTime2018] based on 4930 maximum likelihood gene trees estimated using IQTREE [@minhIQTREE2020]. Divergence dating was estimated using MCMCTree from the PAML package [@yangPAML2007] based on 27 squamate reptile fossil calibrations [@tiatragulAustralia2023]. This phylogeny was used for all phylogenetic comparative analyses.

## Morphological data

We collected body size measurements and vertebral counts from museum specimens examined. To quantify body size and body width ratios, we measured snout-to-vent length (SVL), tail lengths (TL) and midbody diameter (MBD). We measured SVL and TL by stretching the specimen along a ruler to the nearest 0.5 mm. For MBD, we used a digital vernier calliper (ABS Coolant Proof Caliper, Mitutoyo, Japan) to the nearest 0.1 mm. Total body length (TBL) is the sum of SVL + TL. 
<!-- To obtain  mean body shape ratios [@mosimannSize1970] for each species, we divided each MBD by SVL and log-transformed for further analyses. -->

To obtain vertebral counts, we X-ray scanned the specimens at various institutions including: LX-60 (Faxitron, IL, USA) at the University of Adelaide Faculty of Health and Medical Sciences (n=87), Thermo Kevex PXS5-927EA Microfocus source (Waltham, MA, USA) with a LTX-1717 Digital Flat Panel Detector (Kodex Inc, NJ, USA) at the Western Australian Museum (n=129), HeliScan MicroCT system at the National Laboratory for X-Ray Micro Computed Tomography (CTLab) based at the Australian National University (ANU) (n=90), and Queensland X-Ray (n=35).

<!-- Resolution for ANU CTLab is calculated by lenth / pixel * 1000 -->

We used the Fiji distribution of ImageJ2 v.1.54j [@schindelinFiji2012;@ruedenImageJ22017] to digitally visualise X-ray images and counted the number of vertebrae using Cell Counter module. Pre- and post-cloacal vertebrae were distinguished based on vertebral morphology [@szyndlarIllustrated2023]. We calculated vertebrae ratio by dividing the total number of vertebrae by total body length.

Additionally, we quantified intracolumnar length of each vertebra for three species representing 1) narrow-and-wide phenotypes (*A. grypus*), 2) wide-and-robust phenotypes (*A. pinguis*), and 3) intermediate phenotypes (*A. centralis*). Following @sherrattMacroevolution2022 we used the 'multipoint tool' in ImageJ2 to place landmarks medially along the vertebral column at the anterior limit of the centrum of each vertebra. We then calculated the inter-landmark distances by applying the Pythagorean theorem between sequential landmark coordinates in R Statistical Environment v.4.3.1 [@rcoreteamLanguage2023]. Measurements were repeated three times to account for measurement errors.

## Environmental data 
<!-- NEED TO PUT IN THE CORRECT VALUES HERE -->

We used the Australian blindsnake occurrence records presented in Tiatragul et al. (unpublished) including a total of 3898 verified records (min=1, median=23, mean=81, max=525 per species). These records were compiled from museum databases and verified human observations from Atlas of the Living Australia (ALA).

We selected soil bulk density and annual mean temperature, as we are interested in the potential effects these may have on the distribution of species. Soil bulk density measures the penetrability of soil, which is a proxy for the type of microhabitat species is most likely living in as they are all strictly fossorial. Bulk density values > 1.6 g/cm<sup>3</sup> indicate soil that restrict penetration by plant roots, while values <0.5 indicate looser, more organic soil [@mckenzieSoil2002]. Mean annual temperature is a climate proxy and is commonly used to describe niche for squamate reptiles [@coxEvolutionary2015;@pieEvolution2017;@wiensDiversity2013].

From raster files that contained modelled approximations at the resolution of 3 arc-seconds, we extracted soil bulk density values between 0 — 5cm depth from [CSIRO Soil and Landscape Grid of Australia](https://www.clw.csiro.au/aclep/soilandlandscapegrid/index.html), mean annual temperature from WorldClim v.2 [@fickWorldClim2017] for each georeferenced record using the "extract" function in the R package *`raster`* v.3.6-3 [@hijmansRaster2022]. We then calculated the species maximum soil bulk density and species mean annual temperature.

## Statistical analyses

All statistical analyses were performed in R Statistical Environment. To test for evolutionary correlations between the number of vertebrae and body size and proportions, we used phylogenetic generalised least squares (PGLS) regression under a model of Brownian Motion implemented in the R package *`geomorph`* v.4.0.6 [@adamsGeomorph2023]. First, we tested for pleomerism [@lindseyPleomerism1975] across 37 species of *Anilios* blindsnakes. We did this by testing whether the total number of vertebrae positively correlate with body size when regressing the maximum number of total vertebrae against maximum total body length from Tiatragul et al. (unpublished). Next, we test whether the vertebrae to body length ratio is related to the mean vertebrae number and mean aspect ratio, defined by total body length / body width, also from Tiatragul et al. (unpublished) data. Finally, we tested if variation in vertebrae ratio and vertebrae number can be explained by mean annual temperature and soil compactness. In all PGLS analyses, statistical significance was assessed by permutation, using 1000 iterations.

To visualise the relationship between vertebrae number and intracolumnar vertebral length, for each of the representative species, we plotted the mean vertebrae size (from three repeated measurements) against the vertebra number. Each vertebrae profile was overlaid with a fitted 4° polynomial curve.

# Results

### Prevalence of pleomerism among blindsnakes

Total number of vertebrae varies among the 39 species of Australo-Papuan blindsnakes in this study (Fig. \@ref(fig:fig-tree-boxplots)), ranging from 146 in *A. tovelli* to 400 in *A. grypus* (**Table S1**). Aspect ratios ranged from 20.8 in the robust-and-wide *A. pinguis* to 169.8 in the narrow-and-long *A. longissimus* (**Table S2**).

```{r fig-tree-boxplots, fig.cap= 'Interspecific variation in aspect ratio and total number of vertebrae among Australo-Papuan blindsnake and close relatives. The boxplots represent the distribution of data, with vertical inside representing the median. The extent of the box represents the lower (left) and upper (right) quartile, with the whiskers showing 1.5 times the interquartile range. Outliers beyond this range are depicted as individual points. Some species are represented by only the mean.', out.width="165mm", out.height= "129mm"}

knitr::include_graphics(paste0(WD, '/assets/figures/tree-boxplots.jpg'))
```

For the Australian blindsnakes, we found a weakly positive but non-statistically significant relationship between maximum number of total vertebrae and maximum SVL (PGLS, slope=0.28, intercept=256, R^2^=0.02, F=0.74, Z=0.30, P=0.41), indicating an absence of pleomerism in Australian blindsnakes (Fig. \@ref(fig:fig-tbl-vert)). This suggests that maximum length is achieved by elongating each vertebra rather than by adding additional vertebrae, while most species have many more vertebrae than expected for their size. For example, *A. grypus* "West" is an intermediate length (relative to other species) and has the most vertebrae, while *A. torresianus*, which is the longest, has relatively few vertebrae.

### Relationship between vertebrae ratio, number, and body shape
We found that vertebrae ratio is positively correlated with aspect ratio (PGLS, slope=-0.0048, intercept=0.73, R^2^=0.17, F=7.09, Z=2.20, P=0.011), indicating that species that are more narrow-and-long usually have relatively more vertebrae for their length than species that are wide-and-robust (Fig. \@ref(fig:fig-vert-aspect-ratios) and \@ref(fig:fig-vertebral-profile)). We also found that the mean number of vertebrae positively correlated with aspect ratio (PGLS, slope=-1.8, intercept=120.95, R^2^=0.69, F=79.39, Z=5.30, P<0.001).

<!-- Fig. Vertebral counts vs max body length  -->
```{r fig-tbl-vert, fig.cap= 'Maximum vertebral counts plotted against maximum total body length for Australian typhlopids. Trend line represent PGLS regression for Australian species.', out.width="165mm", out.height= "124mm"}

knitr::include_graphics(paste0(WD, '/assets/figures/tbl-vert.jpg'))
```

<!-- Fig: Pleomerism and ratio correlations -->
```{r fig-vert-aspect-ratios, fig.cap= 'Species mean aspect ratios against (A) mean vertebrae ratio and (B) mean number of vertebrae of Australian blindsnakes. Vertebrae ratio is calculated by total vertebrae/total length, and aspect ratio is total body length/midbody diameter. Species with large aspect ratios are narrow-and-long while species with low aspect ratios robust-and-wide.', out.width="124mm", out.height= "165mm"}

knitr::include_graphics(paste0(WD, '/assets/figures/vert-aspect.jpg'))
```

<!-- Fig: Vertebral profile -->
```{r fig-vertebral-profile, fig.cap= 'Relative vertebral length plotted against vertebra number for three example species with fitted 4° polynomial curves overlaid.', out.width="140mm", out.height= "197mm"}

knitr::include_graphics(paste0(WD, '/assets/figures/vertebral-profile.jpg'))
```

### Correlation between vertebrae ratio, number, and ecological variables among Australian blindsnakes

We found a positive relationship between vertebrae ratio and species mean annual temperature (PGLS, slope=0.038, y=0.19, R^2^=0.21, F=9.58, Z=2.32, P=0.009), indicating species that have greater number of vertebrae for body size are generally found in places that are warmer (Fig. \@ref(fig:fig-vert-ecology)A). We also found a significant negative relationship between vertebrae ratio and species maximum soil bulk density (PGLS, slope=-1.00, y=2.52, R^2^=0.12, F=4.8, Z=1.7, P=0.043), indicating species that have greater number of vertebrae for body size are generally found in places that have harder to penetrate soil (Fig. \@ref(fig:fig-vert-ecology)B). When considering raw mean vertebrae number, we found non-statistically significant positive correlations with both mean annual temperature (PGLS, slope=4.13, y=142.45, R^2^=0.76, F=2.87, Z=1.31, P=0.09) and maximum soil bulk density (PGLS, slope=111.69, y=, R^2^=0.04, F=1.61, Z=0.85, P=0.22; Fig. \@ref(fig:fig-vert-ecology)C–D).

<!-- Fig: Ratio and ecological variables -->
```{r fig-vert-ecology, fig.cap= 'Relationship between mean vertebrae ratios and ecological variables. (A) Mean vertebrae against species mean annual temperature. (B) Mean vertebrae ratio against species max soil bulk density. Mean vertebrae ratio is calculated by mean total vertebrae / mean total length for each species. Species mean annual temperature and max soil bulk density values are calculated based on occurrence data. Trend lines represent slopes from PGLS', out.width="165mm", out.height= "117mm"}

knitr::include_graphics(paste0(WD, '/assets/figures/vert-ecology.jpg'))
```

# Discussion

 <!-- The specific questions we addressed are: (1) How prevalent is pleomerism among blindsnakes? (2) Can deviations from pleomerism be explained by body shape in a genus of Australian blindsnakes? (3) Does vertebral ratio vary with mean annual temperature or soil compactness? -->

***[Big summary]***

Results from our study revealed that pleomerism does not apply to Australian typhlopids, indicating a deviation from the positive correlation between body size and vertebrae number reported for other terrestrial snakes and elongated vertebrates [@headDissociation2007;@lindseyPleomerism1975]. Variation in body shape, however, is underpinned by the ratio between vertebrae number to body size. We show that species that are narrow-and-long have proportionally more vertebrae compared to species that are more robust-and-wide. This relationship between the proportion of vertebrae to body shape has been found in colubrids and viperids, where more elongated species tend to have a greater number of vertebrae than stouter ones [@lindellEvolution1994]. Variation in vertebral ratio can also be explained by both mean annual temperature and soil compactness. These results have important implications for understanding the macroevolutionary dynamics of continental-scale radiation of Australian blindsnakes.

***[Discuss why body shape might influence vertebrae ratio or vice-versa]***

Body shape varied with the number of vertebrae in similar ways as previously reported in vipers and colubrids [@lindellEvolution1994]. 
We can speculate that species that are narrow-and-long may be using pre-existing burrows rather than digging their own burrow. Narrow-and-wide phenotypes are common among Leptotyphlopidae, which are thought to use existing burrows rather than dig their own burrow [@herrelBurrowing2021]. In this regard, the Australian typhlopid genus, *Anilios*, is quite exceptional because there is a great variation in body shape (Tiatragul et al. unpublished). Species that evolved more narrow-and-long bodies may be converging on the lifestyle with other scolecophidians that have similar phenotypes (i.e. leptotyphlopids). On the other hand, robust-and-wide phenotypes common among typhlopid species are expected to be able to generate much more force and make their own burrows [@herrelBurrowing2021].

***[Discuss/speculate why vertebrae ratio is correlated with annual mean temperature and soil bulk density]***

Both mean annual temperature and soil bulk density were significant predictors of vertebrae ratio but in different directions. Mean annual temperature has been shown to be a proxy for several ecological factors such as rainfall and vegetation (CITE). However, temperature has also been shown to directly influence multiple axes of reptile development (cite). Vertebral development (somitogenesis) and homeotic (*Hox gene*) regulation may thus be affected by variations in incubation conditions [@martin-del-campoHox2018]. However, we currently lack knowledge about egg-laying behaviour and how temperature affects the development of blindsnake embryos. Common garden experiments across different species of blindsnakes may be able to shed light on whether variation in incubation temperature and consequently, the rate of development may alter the number of vertebrae formation during development. 
<!-- We would expect higher temperature to accelerate development (CITE anolis papers, turtles, etc), so less time for embryonic development.  -->

A previous study on Australian blindsnakes (Tiatragul et al. unpublished) found that robust-and-wide species tend to occupy areas in Australia that have easier-to-penetrate soil compared to narrow-and-long species that tend to occupy areas with more difficult-to-penetrate soil. A greater number of vertebrae in a narrow-and-long body may afford greater flexibility that enables these species to navigate in pre-existing burrows or termite galleries, typical across northern Australia, without exerting force to dig their own burrow. However, more detailed species-level microhabitat observations are required to determine if body shape is related to the environment in which they burrow.

## Miscellaneous 
Traditionally, ventral scale counts have been used as a proxy for vertebral number as they correlate for most alethinophidian snakes [@alexanderPattern1966]. We found that this is (not) the case in blindsnakes. Blindsnakes are generally small enough to fit in an X-ray scanner, even in portable field X-ray machines. Collecting more data on this group of basal snakes can help us understand how vertebral morphology evolved in snakes.

\newpage

## References