### 2024-10-09 manuscript.Rmd
addressed comments from Scott

---
# title: "Dissociation between number of vertebrae and body size in Australian blindsnakes"
title: "Testing the correlates of huge vertebrae number disparity in Australian blindsnakes"
date: "Version `r Sys.Date()`"
author: ""
output: 
  # bookdown::html_document2:
  #   toc: true
  #   number_sections: FALSE
  bookdown::word_document2:
    reference_docx: [./custom-reference-doc.docx]
    number_sections: false
    fig_caption: yes
bibliography: [C:/Users/ST/Documents/repo/anu_thesis_texts/zotero_library.bib]
csl: evolution.csl
link-citations: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE,
                      dpi=300)
# knitr::opts_knit$set(root.dir=normalizePath("..")) #use grandparent folder working directory
# for including graphics we need full path with knitr
WD <- getwd()
```

```{r packages}
library(dplyr); 
library(kableExtra)
```

<!-- **Title** TBD -->

**Authors**: Sarin Tiatragul^1^; Andrew Amey^2^; Scott Hocknull^2^; J. Scott Keogh^1^; Emma Sherratt^3^

^1^Division of Ecology & Evolution, Research School of Biology, The Australian National University, Canberra 2601, ACT, Australia

^2^Queensland Museum, Hendra, QLD, 4011, Australia

^3^School of Biological Sciences, The University of Adelaide, Adelaide, SA 5005, Australia

**Corresponding author**: Sarin Tiatragul (sarin.tiatragul@anu.edu.au)

**CRediT author statement**:

ST: Conceptualisation, Methodology, Formal analysis, Writing - Original draft, reviewing and editing.
AM: Resources, Investigation
SH: Resources, Investigation
SK: Supervision, Funding acquisition, Writing - reviewing and editing.
ES: Conceptualisation, Methodology, Writing - reviewing and editing.

<!-- ST and ES conceptualised the study; ST collected morphology data -->
<!-- ST obtained X-rays from SAM, WAM, and MAGNT. SH and AM provided X-rays Queensland Museum specimens. -->
<!-- ST and ES analysed the data; ST prepared the figures; ST, SK and ES wrote the manuscript. -->

**Acknowledgements**:

We thank museum curators and collection managers from the 
South Australian Museum, Western Australian Museum, Queensland Museum, Australian National Wildlife Collection, Museum and Art Gallery of the Northern Territory, Museums Victoria, and the Florida Museum of Natural History.
SK thanks the Australian Research Council for ongoing support. ST PhD Scholarship was supported by the Australian Government Research Training Program Scholarship at the Australian National University.

**Conflict of Interest statement**
The authors declare no Conflict of Interest.

**Data Availability Statement**
Upon acceptance, associated data will be archived on Zenodo/Dryad.

\newpage

# Abstract

The evolution of the axial skeleton is important in understanding the evolution of body plans in vertebrates. Changes in the vertebral skeleton can influence body form and modes of locomotion. Elongate animals are commonly show a positive correlation between body size and the number of vertebrae, a condition known as pleomerism. However, some groups of vertebrates show a lack of pleomerism, suggesting it is possible to dissociate the production of segment and somatic growth, but what factors influenced this pattern is not well understood. Snakes offer a great model to study axial evolution because the number of vertebrae vary greatly between species and they exhibit a wide range of body size and shape. Here, we investigate the prevalence of pleomerism in a continental radiation of fossorial blindsnakes and test whether body shape and ecology can explain the variation in vertebrae number. We obtained X-ray scans and morphometric data for 42 described species of *Anilios* blindsnakes and evaluated their correlation with each other and ecological variables. We found an absence of pleomerism among Australian blindsnakes and a correlation between vertebrae ratio and body shape, where narrow-and-long species usually possess more vertebrae for their maximum length than wide-and-robust species. Finally, we found that species that have a greater number of vertebrae for their size are generally found in warmer environments and environments with harder-to-penetrate soil substrate. Our results extend our understanding of the axial evolution of basal snakes and the macroevolutionary dynamics of Australian blindsnakes.

Keywords: Scolecophidia, vertebrae, pleomerism, skeleton, axial

\newpage

# Introduction

Evolution of the axial skeleton is fundamental to vertebrate body plan evolution, facilitating body form changes and influencing modes of locomotion in mammals [e.g., @taewcharoenSmall2024;@galisFast2014], reptiles [e.g., @vandammeSpeed2002;@molnarMorphological2023], amphibians [e.g., @bonettEvidence2017;@adlerEvolutionary2022] and fish [e.g., @yamadaAdaptive2009;@forkerBendy2021]. Axial skeleton evolution is often studied by counting the number of vertebrae because the developmental mechanisms of segmentation are well characterised. Essentially vertebrae begin as somites, epithelial blocks of mesoderm, which form rhythmically during embryogenesis where the speed of this process relates to the size and number of somites, to produce what will become the axial skeleton [@gomezControl2008;@dequeantSegmental2008;@gomezDevelopmental2009]. Elongate animals are expected to show a positive correlation between maximum adult body length and number of vertebrae, known as pleomerism [@lindseyPleomerism1975]. While pleomerism was considered the norm [@lindseyPleomerism1975;@wakeComparative1966;@pollyTesting2001], more recent studies have identified exceptions to pleomerism rule in moray eels [@reeceEvolutionary2013], large species of snakes [@headDissociation2007;@lindellEvolution1994], and salamanders and caecilians [@bonettEvidence2017;@renousBody1989;@arntzenVertebral2015].

The lack of pleomerism in some species suggests it is possible to disassociate segment production with body growth, and thus reveals the developmental processes underlying evolutionary patterns. Another clue to disassociation in reptiles can be found in the relationship between the number of ventral scales and precloacal vertebrae – a correlation that is 1:1 in many snakes [@alexanderPattern1966] and so ventral scales have often been used as a proxy for vertebrae number [@lindellEvolution1994;@shineVertebral2000]. However, this 1:1 correlation is absent in file snakes,*Achrochodus*, and some blindsnakes, Typhlopidae, [@alexanderPattern1966;@headDissociation2007], as well as some sea snakes of the genus *Hydrophis* [@vorisDermal1975], suggesting these taxa do not conform to pleomerism.
Snakes are therefore a good model to study axial evolution because they have great diversity in vertebrae number and a wide range of adult body sizes and shapes but without a limb skeleton to confound signals of locomotor adaptation. Recent studies have indicated the number of vertebrae varies predictably with ecological diversification in elapid snakes [@sherrattVertebral2019;@hamptonInterspecific2019;@sherrattMacroevolution2022] and colubrid snakes [@nash-hahnModification2024]. 
The number of vertebrae has also been shown to influence snake locomotion, whereby the number of vertebrae negatively correlates with speed travelled in garter snakes [@kelleyEffects1997]. Despite these recent advances, there have been no study that have quantified the number of vertebrae across all the species in a diverse assemblage to test the prevalence of pleomerism in fossorial snakes.

<!-- [Specific background]-->
Scolecophidians, commonly known in English as "blindsnakes" and "threadsnakes", are characterised by a reduced visual system and highly specialised phenotypes adapted for a strictly fossorial existence [@mirallesMolecular2018;@strongDeconstructing2021;@gowerEyetranscriptome2021]. Scolecophidians comprise five paraphyletic families, including the "core scolecophidians" — a monophyletic clade consisting of Typhlopidae (425 spp), Leptotyphlopidae (143 spp), Gerrhopilidae (29 spp), and Xenotyphlopidae (1 spp) — and the paraphyletic Anomalepididae (20 spp), which is sister to all other "typical snakes" (Alethinophidia) [@mirallesMolecular2018; @burbrinkInterrogating2020; @uetzQuarter2021]. Although all blindsnakes are fossorial, body size and shape vary significantly among species [@tiatragulMorphological2024; @shineNatural1990;@webbPreysize1993;@webbDietary2001;@ehmannFamily1993]. Leptotyphlopids are generally considered more narrow-and-long compared to typhlopids, which are generally wide-and-robust for a given body length [@hedgesLower2008;@webbDietary2001;@webbPreysize1993;@iturriagaNew2021;@tiatragulMorphological2024]. These differences in body size and proportions determine how much push force the snakes can generate when burrowing, which led to the idea that leptyphlopids generally use existing burrows while typhlopids create their own burrow [@herrelBurrowing2021].

The relationship between segmentation and body elongation in scolecophidians is not well understood. However, a study by Head and Polly [-@headDissociation2007] that included 49 species across two genera typhlopids showed that vertebrae number and body size are positively correlated except when some of the largest species were included in the analysis. In particular, the largest species of blindsnakes in the African continent (*Afrotyphlops schlegelii*) have more precloacal vertebrae than expected for their size [@headDissociation2007]. This suggests that elongation in these species is achieved by elongating each vertebra rather than adding additional vertebrae. Compared to other snakes, the vertebral column of fossorial snakes tends to feature a much higher proportion of pre- than post-cloacal segments [@pollyTesting2001], which is reflected in the relatively short tails of blindsnakes [@tiatragulMorphological2024;@szyndlarIllustrated2023]. Therefore, blindsnakes are a good system to test how changes in the number of pre-cloacal vertebrae affects maximum body length.

<!-- Limited studies on scolecophidian axial skeletons compared to alethinophidia means we have yet to appreciate the diversity that these groups have to offer. Given the recent shift in the perception of these earliest diverging clades of snakes (scolecophidians) away from the traditional view that they are "plesiomorphic and morphologically homogenous" [@strongConvergence2022], the scolecophidian axial skeleton also could offer a new perspective on the ecological origin of snakes.  -->
<!-- In another predominantly burrowing limbless clade, caecilian amphibians demonstrate a lack of pleomerism [@renousBody1989].  -->
<!-- Burrowing amphisbaenian lizards and some typhlopid blindsnakes have more ventral scales than vertebrae [@alexanderPattern1966] suggesting pleomerism. -->

The Australian blindsnake radiation, *Anilios*, comprise 47 described species distributed across all major biomes of Australia [@tiatragulAustralia2023;@tiatragulPaleoenvironmental2023]. Species in the genus show considerably diverse morphology from narrow-and-long to robust-and-wide [@tiatragulMorphological2024;@ehmannFamily1993]. However, information about the relationship between body size and vertebrae number is unknown. In this study, we investigate the prevalence of pleomerism among Australian blindsnakes and test hypotheses about how deviations from pleomerism may be explained by different body shapes or other ecological factors among Australian typhlopid snakes. The specific questions we addressed are: (1) How prevalent is pleomerism among Australian blindsnakes? (2) Can deviations from pleomerism be explained by body shape in Australian blindsnakes? (3) Does vertebral ratio vary with their ecology (e.g., soil compactness and mean annual temperature)?

<!-- Specifically, we test two predictions: -->
<!-- (1) narrow-and-long species develop from postnatal ontogenetic changes where vertebral numbers are not significantly different between narrow-and-long species compared to wide-and-robust species.  -->
<!-- (2) Narrow-and-long blindsnakes develop via changes in the embryonic and postnatal stages where narrow-and-long species have greater number of vertebrae than wide-and-robust species.  -->

<!-- To answer these questions we gathered data on the number of vertebrae across XX species of Australian blindsnakes.  -->

# Material and Methods

## Sampling

We examined 341 specimens representing 44 lineages *Anilios* blindsnakes including 42/47 formally described species [@uetzQuarter2021]. We additionally included two close relatives, *Ramphotyphlops multilineatus* and *Acutotyphlops subocularis* [@tiatragulAustralia2023]. Our sample size per species ranged from 1 to 25 specimens (median=7 specimens). We sourced specimens from South Australian Museum (SAMA), Western Australian Museum (WAM), Queensland Museum (QM), Museum and Art Gallery of the Northern Territory (MAGNT), and a private collection. Where possible, we chose relatively large non-fragile specimens from both males and females with well-preserved heads for X-ray radiography. Sex for each specimen were determined by inspecting the gonads or assigned based on relative tail lengths following @tiatragulMorphological2024.

## Phylogeny

We extracted 37 lineages of Australian blindsnakes from the time-calibrated phylogeny of Australian blindsnakes and outgroups presented in @tiatragulAustralia2023. The subset phylogeny represents 35/47 formally described species of _Anilios_ along with two additional genetically distinct lineages: one eastern population of "_A. grypus_" and one eastern population of "_A. ligatus_" [@tiatragulAustralia2023]. This phylogeny was inferred using the shortcut coalescent method in ASTRAL III [@zhangASTRALIIIPolynomialTime2018] based on 4930 maximum likelihood gene trees estimated using IQTREE [@minhIQTREE2020]. Divergence dating was estimated using MCMCTree from the PAML package [@yangPAML2007] based on 27 squamate reptile fossil calibrations [@tiatragulAustralia2023]. This phylogeny was used for all phylogenetic comparative analyses.

## Morphological data

We collected body size measurements and vertebral counts from museum specimens examined. Body measurements including snout-to-vent length (SVL), tail lengths (TL) and midbody diameter (MBD) were obtained from @tiatragulMorphological2024. SVL and TL were measured to the nearest 1 mm on a ruler. MBD was measured using a digital vernier calliper (ABS Coolant Proof Caliper, Mitutoyo, Japan) to the nearest 0.1 mm. Total body length (TBL) is the sum of SVL and TL. To quantify body shape for each specimen, we calculated the aspect ratio [also referred to as elongation ratio; *sensu* @wardConvergent2004], which is the quotient of TBL and MBD.

To obtain vertebral counts, we X-ray scanned the specimens on the following machines and institutions: LX-60 (Faxitron, IL, USA) at the University of Adelaide Faculty of Health and Medical Sciences (n=87), Thermo Kevex PXS5-927EA Microfocus source (Waltham, MA, USA) with a LTX-1717 Digital Flat Panel Detector (Kodex Inc, NJ, USA) at the Western Australian Museum (n=129), HeliScan MicroCT system at the National Laboratory for X-Ray Micro Computed Tomography (CTLab) based at the Australian National University (n=90), and Queensland X-Ray (n=35).

<!-- Resolution for ANU CTLab is calculated by lenth / pixel * 1000 -->

We used the Fiji distribution of ImageJ2 v.1.54j [@schindelinFiji2012;@ruedenImageJ22017] to digitally visualise X-ray images and counted the number of vertebrae using the Cell Counter module. Pre- and post-cloacal vertebrae were distinguished based on vertebral morphology [@szyndlarIllustrated2023]. We calculated vertebrae ratio by dividing the total number of vertebrae by maximum TBL.

Additionally, we quantified intracolumnar length of each vertebra for three species representing (1) narrow-and-wide phenotypes (*A. grypus*), (2) wide-and-robust phenotypes (*A. pinguis*), and (3) intermediate phenotypes (*A. centralis*). Following @sherrattMacroevolution2022 we used the 'multipoint tool' in ImageJ2 to place landmarks medially along the vertebral column at the anterior limit of the centrum of each vertebra. We then calculated the inter-landmark distances by applying the Pythagorean theorem between sequential landmark coordinates in R Statistical Environment v.4.4.0 [@rcoreteamLanguage2024]. Measurements were repeated three times to account for measurement errors.

## Environmental data

We used the Australian blindsnake occurrence records presented in @tiatragulMorphological2024 including a total of 3898 verified records (min=1, median=23, mean=81, max=525 per species). These records were compiled from museum databases and verified human observations from Atlas of the Living Australia (ALA).

We selected soil bulk density and annual mean temperature, as we are interested in the potential effects these may have on the distribution of species. Soil bulk density measures the penetrability of soil, which is a proxy for the type of microhabitat species is most likely living in as they are all strictly fossorial. Mean annual temperature is a climate proxy and is commonly used to describe niche for squamate reptiles [@coxEvolutionary2015;@pieEvolution2017;@wiensDiversity2013]. Temperature also directly influence multiple axes of reptile development [@nobleDevelopmental2018;@boothInfluence2006;@feinerDevelopmental2018], which influences vertebrae development [@martin-del-campoHox2018]. Bulk density values > 1.6 g/cm<sup>3</sup> indicate soil that restrict penetration by plant roots, while values <0.5 indicate looser, more organic soil [@mckenzieSoil2002].

We extracted soil bulk density values between 0 — 5cm depth from [CSIRO Soil and Landscape Grid of Australia](https://www.clw.csiro.au/aclep/soilandlandscapegrid/index.html) and mean annual temperature from WorldClim v.2 [@fickWorldClim2017] from raster files that contained modelled approximations at the resolution of 3 arc-seconds. For each georeferenced record using the "extract" function in the R package *`raster`* v.3.6-3 [@hijmansRaster2022]. We then calculated the species maximum soil bulk density and species mean annual temperature.

## Statistical analyses

All statistical analyses were performed in R Statistical Environment v 4.4.0 [@rcoreteamLanguage2024]. To test for evolutionary correlations between the number of vertebrae and body size and proportions, we used phylogenetic generalised least squares (PGLS) regression under a model of Brownian Motion implemented in the R package *`geomorph`* v.4.0.6 [@adamsGeomorph2023]. First, we tested for pleomerism [@lindseyPleomerism1975] across 37 species of *Anilios* blindsnakes. We did this by testing whether the total number of vertebrae positively correlate with body size when regressing the maximum number of total vertebrae against maximum TBL. Next, we tested whether vertebrae ratio is related to the mean vertebrae number and mean aspect ratio. Finally, we tested if variation in vertebrae ratio and vertebrae number is correlated with maximum soil bulk density and mean annual temperature. In all PGLS analyses, statistical significance was assessed by permutation, using 1000 iterations. To visualise the relationship between vertebrae number and intracolumnar vertebral length, for each of the representative species, we plotted the mean vertebrae size (from three repeated measurements) against the vertebra number. Each vertebrae profile was overlaid with a fitted 4° polynomial curve [e.g., @sherrattMacroevolution2022].

# Results

### Prevalence of pleomerism among blindsnakes

Total number of vertebrae varies greatly among the 39 species of Australo-Papuan blindsnakes in this study (Fig. \@ref(fig:fig-tree-boxplots)), ranging from 146 in *A. tovelli* to 400 in *A. grypus* (**Table S1**). Aspect ratios also vary greatly, ranging from 20.8 in the robust-and-wide *A. pinguis* to 169.8 in the narrow-and-long *A. longissimus* (**Table S2**). Intraspecific variation is also present in both aspect ratio and total number of vertebrae (Fig. \@ref(fig:fig-tree-boxplots)).

```{r fig-tree-boxplots, fig.cap= 'Interspecific variation in aspect ratio and total number of vertebrae among Australo-Papuan blindsnake and close relatives. The boxplots represent the distribution of data, with vertical thick line inside representing the median. The extent of the box represents the lower (left) and upper (right) quartile, with the whiskers showing 1.5 times the interquartile range. Outliers beyond this range are depicted as individual points. Some species are represented by only the mean.', out.width="165mm", out.height= "129mm"}

knitr::include_graphics(paste0(WD, '/assets/figures/tree-boxplots.jpg'))
```

For the Australian blindsnake genus *Anilios*, we found a weakly positive but non-statistically significant relationship between maximum number of total vertebrae and maximum TBL (PGLS, slope=0.28, intercept=256, R^2^=0.02, F=0.74, Z=0.30, P=0.41), indicating an absence of pleomerism in Australian blindsnakes (Fig. \@ref(fig:fig-tbl-vert) and **Fig S1**). This suggests that maximum length is achieved by elongating each vertebra rather than by adding additional vertebrae. Many species have more vertebrae than expected for their size. For example, *A. grypus* "West" is an intermediate length (relative to other species) but has the greatest number of vertebrae, while *A. torresianus*, which has the greatest maximum TBL, has relatively few vertebrae.

### Relationships between vertebrae ratio, number, and aspect ratio
We found that vertebrae ratio is positively correlated with mean aspect ratio (PGLS, slope=-0.0048, intercept=0.73, R^2^=0.17, F=7.09, Z=2.20, P=0.011), indicating that species that are more narrow-and-long usually have relatively more vertebrae for their length than species that are wide-and-robust (Fig. \@ref(fig:fig-vert-aspect-ratios)). We also found that the mean number of vertebrae positively correlated with aspect ratio (PGLS, slope=-1.8, intercept=120.95, R^2^=0.69, F=79.39, Z=5.30, P<0.001).

<!-- Fig. Vertebral counts vs max body length  -->
```{r fig-tbl-vert, fig.cap= 'Maximum vertebral counts plotted against maximum total body length (TBL) for Australian typhlopids. The trend line represents PGLS regression for Australian species. Phylogeny projected into space using maximum likelihood ancestral state estimation. Species not included in the phylogeny are shown in purple. ', out.width="165mm", out.height= "124mm"}

knitr::include_graphics(paste0(WD, '/assets/figures/phylomorpho-tbl-vert.jpg'))
```

<!-- Fig: Pleomerism and ratio correlations -->
```{r fig-vert-aspect-ratios, fig.cap= 'Species mean aspect ratios against (A) mean vertebrae ratio and (B) mean number of vertebrae of Australian blindsnakes. Vertebrae ratio is calculated by total vertebrae/TBL, and aspect ratio is TBL/MBD. (C) Vertebral profile shape of three species representing different vertebrae ratios, and are depicted as a 4° polynomial curve with positive inflection. Species with large aspect ratios are narrow-and-long while species with low aspect ratios robust-and-wide. Phylogeny projected into space using maximum likelihood ancestral state estimation. Species not included in the phylogeny are shown in purple.', out.width="124mm", out.height= "165mm"}

knitr::include_graphics(paste0(WD, '/assets/figures/phylomorpho-vert-aspect-total.jpg'))
# knitr::include_graphics(paste0(WD, '/assets/figures/vert-aspect.jpg')) -->
```

<!-- Fig: Vertebral profile -->
<!-- ```{r fig-vertebral-profile, fig.cap= 'Relative vertebral length plotted against vertebra number for three example species with fitted 4° polynomial curves overlaid.', out.width="140mm", out.height= "197mm"}

knitr::include_graphics(paste0(WD, '/assets/figures/vertebral-profile.jpg'))
``` -->

### Correlation between vertebrae ratio, number, and ecological variables among Australian blindsnakes

We found a positive relationship between vertebrae ratio and species mean annual temperature (PGLS, slope=0.038, y=0.19, R^2^=0.21, F=9.58, Z=2.32, P=0.009), indicating species that have greater number of vertebrae for body size are generally found in places that are warmer (Fig. \@ref(fig:fig-vert-ecology)A). We also found a significant negative relationship between vertebrae ratio and species maximum soil bulk density (PGLS, slope=-1.00, y=2.52, R^2^=0.12, F=4.8, Z=1.7, P=0.043), indicating species that have greater number of vertebrae for body size are generally found in places that have  harder-to-penetrate soil (Fig. \@ref(fig:fig-vert-ecology)B). When considering raw mean vertebrae number, we found non-statistically significant positive correlations with both mean annual temperature (PGLS, slope=4.13, y=142.45, R^2^=0.76, F=2.87, Z=1.31, P=0.09) and maximum soil bulk density (PGLS, slope=111.69, y=, R^2^=0.04, F=1.61, Z=0.85, P=0.22; Fig. \@ref(fig:fig-vert-ecology)C–D).

<!-- Fig: Ratio and ecological variables -->
```{r fig-vert-ecology, fig.cap= 'Phylomorphospaces showing relationship between mean vertebrae ratios and mean vertebrae number against ecological variables. Mean vertebrae ratio against (A) species mean annual temperature and (B) species max soil bulk density. Mean vertebrae ratio is calculated by mean total vertebrae / mean TBL for each species. Mean vertebrae number against (C) species mean annual temperature and (D) species max soil bulk density. Species mean annual temperature and max soil bulk density values are calculated based on occurrence data. Trend lines represent slopes from PGLS', out.width="165mm", out.height= "117mm"}

knitr::include_graphics(paste0(WD, '/assets/figures/phylomorpho-vert-ecology.jpg'))
```

# Discussion

Our study investigated the presence of pleomerism in Australian blindsnakes and tested whether deviations from pleomerism may be explained by body shape or ecological factors. Our results revealed that pleomerism does not generally apply to Australian typhlopids, indicating a deviation from the positive correlation between maximum body length and vertebrae number reported for other terrestrial snakes and elongated vertebrates [@headDissociation2007;@lindseyPleomerism1975]. Variation in body shape, however, is related to vertebrae ratio. We show that species that are narrow-and-long have proportionally more vertebrae compared to species that are more robust-and-wide. We also found that vertebral ratio correlates positively with mean annual temperature and negatively with soil compactness. These results have important implications for understanding the axial evolution of basal snakes and the macroevolutionary dynamics of Australian blindsnakes.

### Absence of pleomerism in Australian blindsnakes

We found a dissociation between vertebrae number and maximum adult body length in Australian blindsnake. While pleomerism has been documented in most snake groups [@lindellEvolution1994], exceptionally large species ("giant taxa") within boids, pythonids, and blindsnakes such as *Afrotyphlops schlegelii* have fewer than expected pre-cloacal vertebrae for their size [@headDissociation2007]. In line with these observations, some of the largest species in our dataset (*A. torresianus* and *A. pinguis*) possess lower than average number of vertebrae (Fig. \@ref(fig:fig-tbl-vert)). The absence of pleomerism suggest that the large size attained by these "giant taxa" may be due to modification of post-somitogenetic somatic growth [@headDissociation2007]. In other words, species with larger body size do not simply have more vertebrae segments, rather the size of each vertebral segment has been elongated during growth. Several mechanisms have been suggested for how giant amniotes achieve their large size including prolonging their growth [@ericksonHow1999] and increasing the rate of growth [@ericksonGigantism2004]. Unfortunately, without life-history data for blindsnakes, we cannot test if these mechanisms may be responsible. However, elongation may also be achieved in other ways.
The regionalisation of vertebrae, where distinct regions along the axial column exhibit morphological diversification, and evolve relatively independently, may explain the lack of pleomerism in blindsnakes. For example, the absence of pleomerism in *Hydrophis* sea snakes has been attributed to the increase in the number of smaller vertebrae in the forebody region compared to larger vertebrae in the hindbody region [@sherrattMacroevolution2022;@sherrattVertebral2019]. Similarly, colubrids have also been shown to have different regionalisation patterns in the vertebral column that contributes to elongation in different ways [@nash-hahnModification2024]. 
The position of the heart is a key marker for delimiting the regions of the snake axial column [@hamptonHeart2022;@hamptonRegionalization2023]. Future research should investigate the regionalisation of the vertebral column in blindsnakes as this will provide important insights to understanding the lack of pleomerism.

The lack of pleomerism among burrowing vertebrates is not uncommon. @lindellEvolution1994 found that among Australian elapids and African snakes, burrowing species had fewer vertebrae than expected from their body length compared to terrestrial and arboreal species. The dissociation between vertebrae number and body size has also been documented in caecilians, a primarily fossorial order of amphibians. In caecilians, @renousBody1989 found that members of the same genus will vary in size but generally share similar number of vertebrae, indicating a lack of pleomerism except for in a few genera such as *Oscaecilia*. Another group of highly specialised burrowing reptile is the amphisbaenians [@gansCharacteristics1978;@gansAmphisbaeniansreptiles1969]. While several studies have documented the number of vertebrae in some species [@bergmannVertebral2012;@gansStudies1960;@gansStudies1971;@gansCharacteristics1978;@alexanderPattern1966], no study has investigated the correlation between vertebrae number and body size in amphisbaenians. However, they have been shown to deviate from the expected 1:1 ratio of ventral scale number to vertebrae [@alexanderPattern1966], suggesting a lack of pleomerism. We recommend future research to comprehensively investigate the prevalence of pleomerism among burrowing limbless and limb-reduced vertebrates to improve our understanding of the evolution of axial system.

### Correlation between vertebrae ratio, number, and aspect ratio

Aspect ratio (body shape) varied with the number of vertebrae in similar ways as previously reported in fish and other snakes (e.g. vipers and colubrids), where species with a more elongated body had relatively more vertebrae than stouter ones [@lindseyPleomerism1975;@lindellEvolution1994;@wardAxial2010]. Two possible explanations have been proposed by @lindellEvolution1994 for snakes including adaptations for locomotion and prey consumption, which are mirrored in the interpretations from elongate fish [@wardAxial2010]. Regarding locomotion, @poughSpecializations1983 showed that in snakes, longer body segments may be advantageous for elongated species that usually move faster, whereas shorter segments are favoured for stouter species that move more slowly. In blindsnakes, however, movement speed may not be as important as how much push force they can generate when burrowing. Snakes with relatively wider bodies are generally able to produce greater push force for their body size [@herrelBurrowing2021], while species with narrower bodies will be able to use pre-existing small crevices and galleries without needing to expend any energy. In regards to prey consumption, stout species with large heads (i.e. vipers) tend to consume large bulky prey, which could be facilitated by having a stout body as this minimises disruption of physiological functions [@poughSpecializations1983;@lindellEvolution1994;@moonFeeding2019]. Although Australian blindsnakes primarily prey on pupae and larvae of ants and termites, there is some variation in the size of prey consumed [@webbDietary1993]. In particular, heavy-bodied blindsnake species (e.g. *A. pinguis*, *A. proximus*, and *A. ligatus*) fed on larger species of ants compared to more slender-bodied blindsnakes species (e.g. *A. hamatus*, *A. bituberculatus*, *A. waitii*, *A. weidii*) [@webbDietary1993].

<!-- ** Discuss how longer body allow more force by elongating muscles @navasMorphological2004 ** -->

### Correlations between vertebrae ratio, number, and ecological variables
<!-- Discuss/speculate why vertebrae ratio is correlated with annual mean temperature and soil bulk density]*** -->

Both mean annual temperature and soil bulk density were significant predictors of vertebrae ratio but not for the raw number of vertebrae. While non-statistically significant, we observed a slightly positive relationship between raw number of vertebrae and mean annual temperature. Previous studies across different species of rattlesnakes [@klauberRattlesnakes1982;@klauberCorrelation1941] and between populations of garter snakes (*Thamnophis elegans*) [@arnoldQuantitative1988] both reported a similar pattern where species that occupy warmer and dryer environments had more vertebrae than those living in colder and humid coastal areas. While the temperature embryos are exposed to during development has been shown to affect vertebral numbers in snakes [@foxMorphological1961;@foxEffect1948;@osgoodEffects1978], @arnoldQuantitative1988 suggested three reasons it is unlikely the explanation for the regional difference in the number of vertebrae observed in *T. elegans*. First, the observed difference in the number of vertebrae between populations is much greater than the effects of temperature produced lab experiments [@arnoldQuantitative1988;@osgoodEffects1978]. Second, the number of vertebrae of the F1 hybrid is intermediate between inland and coastal parents [@arnoldQuantitative1988]. Finally, females can select microhabitats that buffer their developing embryos against extreme climatic conditions [@arnoldQuantitative1988;@brownMaternal2004;@burgerNest1986;@peet-pareFemale2012]. Although we currently lack knowledge about egg-laying behaviour and how temperature affects the development of blindsnake embryos, we suspect that temperature effects alone are unlikely to explain the interspecific variation in vertebrae number.

In a related study, @tiatragulMorphological2024 found that Australian blindsnake species with smaller absolute body size generally occupy areas that are warmer regions of Australia than while larger species tend to occupy cooler regions. Here, we observe that vertebrae ratio (raw vertebrae number/SVL) is positively correlated with temperature, indicating species that have more vertebrae than expected from their body size are found in warmer places. In his work on rattlesnakes, @klauberRattlesnakes1982  noted that smaller-sized subspecies often retain the same number of vertebrae as their ancestors, which result in the smaller species having relatively greater vertebrae ratio. An explanation for this pattern is that size is often under greater selection than vertebral number, and there is negligible disadvantage for a snake in having too many vertebrae for their body size [@lindellEvolution1994]. This makes sense when we also consider the negative relationship between vertebrae ratio and maximum soil bulk density. A greater number of (shorter) vertebrae for body size may afford greater flexibility [@jayneKinematics1986;@bergmannVertebral2012;@johnsonAdaptive1955] that enables these species to navigate in pre-existing burrows or termite galleries, typical across northern Australia, without exerting force to dig their own burrow. @tiatragulMorphological2024 also found species that occupy areas with more difficult-to-penetrate soil are usually narrow-and-long rather than robust-and-wide species that occupy areas with easier-to-penetrate soil. Narrow-and-wide phenotypes are common among leptotyphlopids, which are thought to use existing burrows rather than digging their own [@herrelBurrowing2021]. In this regard, the Australian typhlopid genus *Anilios* is quite exceptional because there is great variation in body shape [@tiatragulMorphological2024]. Species that evolved more narrow-and-long bodies may be converging on the behaviour with other scolecophidians that have similar phenotypes (i.e., leptotyphlopids). However, more detailed species-level microhabitat observations are required to determine if these species are actually taking advantage of these environments in the ways we speculate.

<!-- ## Miscellaneous  -->
<!-- Traditionally, ventral scale counts have been used as a proxy for vertebral number as they correlate for most alethinophidian snakes [@alexanderPattern1966]. We found that this is not the case in blindsnakes.  -->

## Conclusion

Compared to other genera of blindsnakes, the Australo-Papuan genus *Anilios* show exceptional diversity in body size and shape [@tiatragulMorphological2024;@herrelBurrowing2021], which make them a good group to study evolution of the axial skeleton. We found a lack of pleomerism in the genus, with exceptionally large species possessing a smaller number of vertebrae than expected for their body size. Our findings indicate that body size and shape evolution in this clade has been facilitated by changes to development, whereby the production of vertebral segments during early development is dissociated from somatic growth, likely driven by ecological adaptations. This study adds to a growing body of literature on the axial skeleton of snakes, a field that has largely been neglected.

\newpage

## References